Testing atlantisom: what if we interpolate the weight at stage output to true age?

### 2019-05-16 InterpolateWeightAtAge.Rmd
working on function line 232

---
title: "Testing atlantisom: visualize weight at (st)age output. How should we interplolate this for true ages?"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
bibliography: "packages.bib"
---

@@

Now we delve into the calculation of true age classes and weight-at-age from standard Atlantis output age class, which is generally in 10 stages of maxage/10 true age classes for all but the newest Atlantis models. Note that modern Atlantis models can output true age classes in two output files (from the wiki):

@@

However, The CCA model does not produce this output, and other legacy models may not, so we still need the `atlantisom` functions `calc_stage2age` and the associated `calc_Z`.  This function will be tested. However, it produces numbers at age but not weight. Weight-at-true-age is needed as an input to stock assessment models using empirical weight at age. See discussion [here](https://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html).

Therefore, our first step is understanding what the 10 class weight at age (output as `mulength` from `calc_age2length`) looks like for a variety of species.

@@

## Simulate a survey part 5: weight at st(age) census

Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/StdSurvLengthCompTest.html).

We will apply examples here to just a few NOBA species to examine within-year variability:
    - Cod "North_atl_cod", likely a test assessment species
    - Herring "Norwegian_ssh", likely a test assessment species
    - Greenland halibut "Green_halibut", which grows to a large size.

@@

Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age composition (for just one species in this case). The `sample_fish` applies the median for aggregation and does not apply multinomial sampling if `sample=FALSE` in the function call.

@@

We should get the upper end of Greenland halibut with a 200cm max length bin. That shouldnt matter for weight at age. This is all timesteps for each species by agecl.

```{r vis-wtageclass-test}

wageplot <- ggplot(length_census$muweight, aes(agecl, atoutput)) +
  geom_point(aes(colour = time)) +
  theme_tufte() +
  theme(legend.position = "bottom") +
  scale_x_discrete(limits=c(1:10)) +
  xlab("age class") +
  ylab("average individual weight (g)") +
  labs(subtitle = paste(scenario.name))

How much does weight at age class vary over time? Here are a couple of annual cycles:

```{r vis-wtageclass-test2}

wtageyr1 <- length_census$muweight %>%
  filter(time<5)

wageplot <- ggplot(wtageyr1, aes(agecl, atoutput)) +
  geom_line(aes(colour = factor(time))) +
  theme_tufte() +
  theme(legend.position = "bottom") +
  scale_x_discrete(limits=c(1:10)) +
  xlab("age class") +
  ylab("average individual weight (g)") +
  labs(subtitle = paste0(scenario.name, " initial year"))

```{r vis-wtageclass-test3}

wtageyrmid <- length_census$muweight %>%
  filter(between(time, 50,54))

wageplot <- ggplot(wtageyrmid, aes(agecl, atoutput)) +
  geom_line(aes(colour = factor(time))) +
  theme_tufte() +
  theme(legend.position = "bottom") +
  scale_x_discrete(limits=c(1:10)) +
  xlab("age class") +
  ylab("average individual weight (g)") +
  labs(subtitle = paste0(scenario.name, " year 10"))

```{r vis-wtageclass-test4}

wtageyr20 <- length_census$muweight %>%
  filter(between(time, 100,104))

wageplot <- ggplot(wtageyr20, aes(agecl, atoutput)) +
  geom_line(aes(colour = factor(time))) +
  theme_tufte() +
  theme(legend.position = "bottom") +
  scale_x_discrete(limits=c(1:10)) +
  xlab("age class") +
  ylab("average individual weight (g)") +
  labs(subtitle = paste0(scenario.name, " year 20"))

There are some interesting jumps between age classes. Most pronounced for herring. All of these species would be split into two true ages per age class. Interpolation could be interesting.

@@

@@

@@

How do these three look? Herring had a rough period from model years 15-30 or so.

Now for some CCA groups with multiple true ages per stage.

Pacific hake, 2 true ages per class:  
```{r plotCCAhake}
censuswt <- length_censussurvsamp$muweight %>%
  filter(species == "Mesopel_M_Fish")

@@

Bocaccio rockfish, 5 true ages per age class:
```{r plotCCAbocaccio}
censuswt <- length_censussurvsamp$muweight %>%
  filter(species == "Bocaccio_rockfish")

@@

Yelloweye rockfish, most extreme at 12 true ages per age class:
```{r plotCCAYelloweye}
censuswt <- length_censussurvsamp$muweight %>%
  filter(species == "Yelloweye_rockfish")

@@

I'm not sure any of this helps us decide how to interpolate, but there is clearly contrast in weight at age (class) in both models and example runs.

>>>>

---
title: "Testing atlantisom: what if we interpolate the weight at stage output to true age?"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
bibliography: "packages.bib"
---

@@

Now we delve into the calculation of true age classes and weight-at-age from standard Atlantis output age class, which is generally in 10 stages of maxage/10 true age classes for all but the newest Atlantis models. Weight-at-true-age is needed as an input to stock assessment models using empirical weight at age. See discussion [here](https://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html). We will test the `atlantisom` functions `calc_stage2age` and the associated `calc_Z` that estimate true numbers at age from numbers in output age class separately. However, `calc_stage2age` produces numbers at age but not weight.

We already have 10 class weight at age (output as `mulength` from `calc_age2length`). We visualized weight at stage for selected NOBA and CCA species [here](https://sgaichas.github.io/poseidon-dev/TestWeightAtStage.html). In this page we will try the simplest interpolation to true age-based weight at age possible, and see if that will work for us.

Note that modern Atlantis models can output true age classes in two output files (from the wiki):

@@

@@

## Simulate a survey part 5: interpolated weight at true age census

Full methods to get the census weight at age by Atlantis output age class are explained [here](https://sgaichas.github.io/poseidon-dev/TrueLengthCompTest.html).

We will apply interpolation examples here to just a few NOBA species to examine within-year variability:
    - Cod "North_atl_cod", likely a test assessment species
    - Herring "Norwegian_ssh", likely a test assessment species
    - Greenland halibut "Green_halibut", which grows to a large size.

@@

Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age composition (for just a few species in this case). The `sample_fish` applies the median for aggregation and does not apply multinomial sampling if `sample=FALSE` in the function call.

@@

We should get the upper end of Greenland halibut with a 200cm max length bin. That shouldnt matter for weight at age. All of these species would be split into two true ages per age class. Interpolation could be interesting.

@@

If we were to do simple interpolation on this output, we can approximate weight at true age class:

```{r interpolate-wtage}
# a function to go into atlantisom, sketched here, reuse calc_stage2age as possible
calc_wtTrueage <- function(wtagecl_data, fgs.file) {
  # get number of ages per agecl from fgs.file
  # match species in wtagecl_data to those species in the fgs file
  fgs <- load_fgs(fgs.file)
  species.code <- fgs$Code
  turnedon <- fgs[fgs$IsTurnedOn > 0, ]
  names <- turnedon$Code

  # Figure out the groups that have multiple ages classes in each stage (or
  # cohort)
  multiple_ages <- turnedon[turnedon$NumAgeClassSize>1, c(1,4,10)]
  num_multi_age <- dim(multiple_ages)[1]

  ntimesteps <- length(unique(wtagecl_data$time))

  # find weight increment for a timestep
  # muweight for agecl+1 - muweight for agecl

  # divide weight increment by number of ages per agecl (could decay with increasing age, but how different from equal increment?)

  # recursive: add increment to agecl until hitting agecl+1

  # reunumber agecl 1 to 10*number of ages per agecl

  # output full set of muweight at newly created agecl

}
```

Plot interpolated results for NOBA species:

```{r interp-wtage-test1}

@@

```{r apply-inter-pwtage-CCA}

Plots for species with 1 true age per agecl should not be different:

@@

How do these three look? Identical I hope.

Now for some CCA groups with multiple true ages interpolated per stage.

Pacific hake, 2 true ages per class (20 total):  
```{r plotCCAhake}
censuswt <- length_censussurvsamp$muweight %>%
  filter(species == "Mesopel_M_Fish")

@@

Bocaccio rockfish, 5 true ages per age class (50 total):
```{r plotCCAbocaccio}
censuswt <- length_censussurvsamp$muweight %>%
  filter(species == "Bocaccio_rockfish")

@@

Yelloweye rockfish, most extreme at 12 true ages per age class (120 total):
```{r plotCCAYelloweye}
censuswt <- length_censussurvsamp$muweight %>%
  filter(species == "Yelloweye_rockfish")

@@

Yelloweye should be a good test...
